---
title: "Testing for intraspecific size number tradeoff"
author: "Claire Smith"
date: "2023-06-28"
output: github_document
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

Investigating the relationship between pollen size and number within wind-pollinated species using linear regression: pollen number ~ pollen volume

```{r initialize, message=F}
library(tidyverse)
library(ggplot2)
library(lme4)
library(lmerTest)
source("theme_cs.R")
```

```{r read in data}
prodfull <- read.csv("processed-data/size-prod-norep.csv", stringsAsFactors = T) # has individual-level pollen size and number data
# head(prodfull)
# str(prodfull)
# summary(prodfull)
```

First I'll take a look at the data and make any adjustments that I want to carry throughout the whole script. In later plots I identified two obvious outliers -- I'll remove these. 

```{r check data}
# head(prodfull)
# summary(prodfull)
prod <- prodfull %>% 
  filter(!is.na(Avg_diam) & !is.na(Avg_pol_anth)) %>% # want only entries with both size and prod values
  # remove outliers
  filter(!(Species=="Phleum pratense"&Ind=="C11")) %>% 
  filter(!(Species=="Elymus innovatus"&Ind=="EI01"))
```

First I'll take a look at a summary of the pollen size and number data - means, standard deviations, and CVs so that I can compare how variable species area relative to each other. I'll remove any species with less than 5 points. 

```{r summary of size number data}
prod_summary <- prod  %>% group_by(Species, source, Sex_sys) %>% 
  summarize(mean_diam = mean(Avg_diam, na.rm=T), 
            sd_diam = sd(Avg_diam, na.rm=T),
            cv_diam = sd_diam/mean_diam,
            mean_polanth = mean(Avg_pol_anth, na.rm=T),
            sd_polanth = sd(Avg_pol_anth, na.rm=T),
            cv_polanth = sd_polanth/mean_polanth,
            n=n()) %>%  
  filter(n>=5) %>% 
  droplevels() %>% 
  arrange(Sex_sys, Species) %>% 
  as.data.frame()

# This will be made into a table - I'll add columns with mean +/- sd for that future table
prod_summary$anth_msd <- paste0(round(prod_summary$mean_polanth,0)," \U00B1 ",round(prod_summary$sd_polanth,0))
prod_summary$diam_msd <- paste0(round(prod_summary$mean_diam,0)," \U00B1 ",round(prod_summary$sd_diam,0))
```

I'll save this summary data to a file.

```{r save summary data to file}
prod_summary_filt <- prod_summary 
#write it to a file I will make a table with 
write.csv(prod_summary_filt, "processed-data/sizeprod-summary.csv", row.names = F, fileEncoding = "UTF-8")
```

```{r t test variance in pollen size vs number}
# how variable is pollen size/num across species? 
# Pollen size seems much less variable - I'll run a t test of mean cv for diameter vs mean cv for pollen per anther to determine if pollen size is significantly less variable than pollen production
# mean(prod_summary$cv_diam)
# sd(prod_summary$cv_diam)
# mean(prod_summary$cv_polanth)
# sd(prod_summary$cv_polanth)
cvtest <- prod_summary %>% select(cv_diam, cv_polanth)
t.test(x=cvtest$cv_polanth, y=cvtest$cv_diam, paired=T) # t test across all species
```

Pollen diameter is less variable than pollen size. The CV for pollen diameter is significantly less than the CV for pollen production per anther. 

Now I want to build a model for each species. First I'll go through each species and look at the distributions of pollen size and number to see how symmetric they are and if there are any obvious outliers. 

```{r prepare data for models}
#nest data by species to make it easier to handle
by_spp <- prod %>% group_by(Species, Sex_sys) %>% 
  nest() %>% 
  arrange(Sex_sys, Species) %>% 
  # Keep only species with at least 10 entries %>% 
  mutate(nind = purrr::map(data, nrow)) %>% 
  filter(nind>=10)
# by_spp # 23 species remaining

#inspect distributions of mean pollen diam and pollen production for each species - are they symmetric? any obvious outliers? 
#define function for this
checkfun <- function(df,i){
  spec = as.character(df[i,]$Species)
  n=length(df[i,]$data[[1]]$Avg_diam)
  hist(df[i,]$data[[1]]$Avg_diam, main=paste(spec, "n=",n), xlab="Mean pollen diameter")
  hist(log(df[i,]$data[[1]]$Avg_diam), main=paste(spec, "n=",n), xlab="Log mean pollen diameter")
  hist(df[i,]$data[[1]]$Avg_pol_anth, main=paste(spec, "n=",n), xlab="Mean pollen per anther")
  hist(log(df[i,]$data[[1]]$Avg_pol_anth), main=paste(spec, "n=",n), xlab="Log mean pollen per anther")
}

# checkfun(by_spp,2)
```

I'll plot all the data as pollen production vs size to get an idea of what it looks like, and if there's any clear outliers or potential grouping in the data. 

```{r all tradeoff plots, dpi=200, fig.dim=c(16,14)}
prod %>% 
  filter(Species %in% by_spp$Species) %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  facet_wrap(~Species, scales = "free") + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans") + 
  theme(strip.background = element_rect(linewidth = NULL,
                                        linetype = NULL,
                                        colour = "white"))
```

Now I'll run linear models predicting pollen production by pollen size, and pull out the ones with a significant relationship. 

```{r, linear regression models, cols.print=14}
#define function to run model 
spp_model <- function(df) {
  lm(Avg_pol_anth ~ Avg_diam, data = df)
}
#run models and add model to by_spp dataframe
by_spp <- by_spp %>% 
  mutate(model = purrr::map(data, spp_model)) %>% 
  arrange(Sex_sys, Species) # arrange in order of sex sys, then alphabetical
#get model summary using glance() from "broom"
glance <- by_spp %>% 
  mutate(glance = purrr::map(model, broom::glance)) %>% 
  unnest(glance, .drop=T)
# put this in a viewable format (remove nested data and model info)
glancetab <- glance %>% select(-c(data, model)) %>% 
  arrange(Sex_sys, Species)
#save this in a new table without the ugly list elements, which I can print to a file
glancetab_write <- glance %>% select(-c(data, model)) %>% 
  arrange(Sex_sys, Species) %>% apply(2, as.character) # coercing all columns to characters so that I can write this to a file
write.csv(glancetab_write, "processed-data/tradeoff-reg-model-sum.csv", row.names = F)
#take a look
print(glancetab, n=Inf)

#similar - but using broom::tidy(), gives me model coefficients
tidy <- by_spp %>% 
  mutate(tidy = purrr::map(model, broom::tidy)) %>% 
  unnest(tidy, .drop=T)
tidytab <- tidy %>% select(-c(data, model)) %>% 
  arrange(Sex_sys, Species)
print(tidytab, n=46)
tidytab_write <- tidy %>% select(-c(data, model)) %>% 
  arrange(Sex_sys, Species) %>% apply(2, as.character)
write.csv(tidytab_write, "processed-data/tradeoff-reg-model-coefs.csv", row.names = F)
#take a look
# print(tidytab, n=Inf)

#take a look at r.squared values
# glance %>%
#   ggplot(aes(Species, r.squared)) +
#   geom_point() +
#   theme(axis.text = element_text(angle=90))

```

```{r}
#take only species with significant relationships to plot with linear model overtop
sigsub <- glance %>% filter(p.value<0.05) #7 species
#Plot relationship between size and number for the 7 species with a significant relationship
# sigsub$Species
```

7 species have significant relationships between pollen size and number: Bromus inermis, Elymus innovatus, Festuca campestris, Festuca pratensis, Koeleria cristata, Plantago lanceolata, and Amaranthus retroflexus. 

```{r Bromus inermis tradeoff, dpi=200}
# Bromus inermis
sigsub[which(sigsub$Species=="Bromus inermis"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Elymus innovatus tradeoff, dpi=200}
# Elymus innovatus
sigsub[which(sigsub$Species=="Elymus innovatus"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Festuca campestris tradeoff, dpi=200}
#Festuca campestris
sigsub[which(sigsub$Species=="Festuca campestris"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Festuca pratensis tradeoff, dpi=200}
# Festuca pratensis
sigsub[which(sigsub$Species=="Festuca pratensis"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Koeleria cristata tradeoff, dpi=200}
# Koeleria cristata
sigsub[which(sigsub$Species=="Koeleria cristata"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Plantago lanceolata tradeoff, dpi=200}
# Plantago lanceolata
sigsub[which(sigsub$Species=="Plantago lanceolata"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Amaranthus retroflexus tradeoff, dpi=200}
#Amaranthus retroflexus
sigsub[which(sigsub$Species=="Amaranthus retroflexus"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")

```

Found a significantly negative relationship between pollen volume and number (i.e. size-number trade-off) in 6/15 species (only taking species with at least 15 entries for pollen size/number). I found a significantly positive relationship between size and number in 1 species (Leymus innovatus). 

I'll also run another set of plots with "Date" as a grouping variable.

```{r Bromus inermis tradeoff date}
# Bromus inermis
sigsub[which(sigsub$Species=="Bromus inermis"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Elymus innovatus tradeoff date}
# Elymus innovatus
sigsub[which(sigsub$Species=="Elymus innovatus"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Festuca campestris tradeoff date}
#Festuca campestris
sigsub[which(sigsub$Species=="Festuca campestris"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Festuca pratensis tradeoff date}
# Festuca pratensis
sigsub[which(sigsub$Species=="Festuca pratensis"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Koeleria cristata tradeoff date}
# Koeleria cristata
sigsub[which(sigsub$Species=="Koeleria cristata"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Plantago lanceolata tradeoff date}
# Plantago lanceolata
sigsub[which(sigsub$Species=="Plantago lanceolata"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")
```

```{r Amaranthus retroflexus tradeoff date}
#Amaranthus retroflexus
sigsub[which(sigsub$Species=="Amaranthus retroflexus"),]$data %>% 
  as.data.frame() %>% 
  ggplot(aes(x=Avg_diam, y=Avg_pol_anth, colour=Date)) + 
  geom_point() + 
  geom_smooth(method="lm", se=F) + 
  scale_y_continuous(name="Pollen production per anther") + 
  scale_x_continuous(name=expression(paste("Pollen diameter (", mu, "m",")"))) + 
  theme_cs(font = "sans")

```


