---
title: "Pollination efficiency"
author: "Claire Smith"
date: "2023-06-23"
output: github_document
editor_options: 
  chunk_output_type: console
---

Calculating pollen transfer efficiency - percentage of grains captured per stigma to grains produced per anther. 

Do plants that produce more pollen also capture more pollen? And does this relationship change with sex system (do dioecious species have lower efficiencies because they cannot capture self pollen?)

```{r initialize, message=F}
## Load packages
library(tidyverse)
library(ggplot2)

## Source files
# custom
source("theme_cs.R")
# data cleaning
source("clean-stig-01.R")
source("clean-stig-02.R")
source("clean-anthers-04.R") # calls 1, 2, 3 within itself
```


```{r read in data}
# Stigmatic pollen capture
pc.dat <- read.csv("processed-data/stig-no-rep-spp.csv", stringsAsFactors = T) 
# head(pc.dat)
# str(pc.dat)
#Fix a typo in stigma data
pc.dat$Species <- gsub("Schizacne purpurascens", "Schizachne purpurascens", pc.dat$Species)
# Pollen production (and size)
sizenum.dat <- read.csv("processed-data/size-prod-all.csv", stringsAsFactors = T)
# head(sizenum.dat)
# str(sizenum.dat)
```

```{r process data}
# The stigma and anther data is at the per-individual level right now - summarize it so that it's at the per-species level
# # ** go back and do a satterthwaite pooled sd instead of just lumping it all together - for now just combine all by species
# stig_sum_ind <- pc.dat %>% 
#   group_by(Species, Date, Site, Plant) %>% 
#   summarize(Avg_pollen_ind = mean(Flw_pollen, na.rm=T), # within-individual avg pollen receipt
#             Sd_pollen_ind = sd(Flw_pollen, na.rm=T), # within-individual sd pollen receipt
#             N_flw = n()) %>% 
#   # Now ungroup to find total number of individuals sampled in each sp
#   group_by(Species) %>% 
#   mutate(N_ind=n()) %>% 
#   ungroup()
# 
# # Pool together to get per-species mean pollen receipt and sd pollen receipt
# stig_sum_spp <- stig_sum_ind %>% 
#   group_by(Species) %>% 
#   summarise(Avg_pollen_sp = mean(Avg_pollen_ind, na.rm=T),
#             sd_num = sum((N_flw-1)*Sd_pollen_ind^2),
#             sd_denom = sum(N_flw) - N_ind,
#             Sd_pollen_sp_pooled = sd_num/sd_denom)

stig_spp <- pc.dat %>% 
  group_by(Species, Sex_sys, source) %>%
  summarise(Avg_stig_sp = mean(Flw_pollen, na.rm=T), 
            Sd_stig_sp = sd(Flw_pollen, na.rm=T),
            n_stig = n(),
            SE_stig_sp = Sd_stig_sp/sqrt(n_stig))
# View(stig_spp)
prod_spp <- sizenum.dat %>% 
  group_by(Species, Sex_sys, source) %>% 
  summarize(Avg_prod_anth_sp = mean(Avg_pol_anth, na.rm=T),
            Sd_prod_anth_sp = sd(Avg_pol_anth, na.rm=T),
            n_prod_anth = n(),
            SE_prod_anth_sp = Sd_prod_anth_sp/n_prod_anth)
# View(prod_spp)
# Join data for plotting - use inner_join() because not all species have both capture and prod data
# Keep JF data for pollen production when species overlap - I trust the automated counting with the
# particle counter more than my subsampling counting. It counts all the particles that go through it
# vs I only count a fraction of the total pollen and assume it's uniformly distributed within the sample.
prod_spp$Species[which(prod_spp$source == "CS2021")]


prod_spp_filt <- prod_spp %>% filter(source != "CS2021" |  Species == "Amaranthus retroflexus")

pdat <- inner_join(stig_spp, prod_spp_filt, by = c("Species", "Sex_sys"))
anti_pdat <- anti_join(stig_spp, prod_spp_filt, by = c("Species"))
anti_pdat2 <- anti_join(prod_spp_filt, stig_spp, by = c("Species"))
# head(pdat)
# summary(pdat)
# View(pdat)
```

2 dioecious species, 8 monoecious species, 9 hermaphroditic species 

```{r}
# Re-order levels in sex system column 
pdat$Sex_sys <- as.character(pdat$Sex_sys)
pdat$Sex_sys <- factor(pdat$Sex_sys, levels=c("dioecious", "monoecious", "hermaphroditic"))

# Arrange data so that species are grouped by sex system and ordered alphabetically
pdat <- arrange(pdat, Sex_sys, Species)
pdat$Species <- factor(pdat$Species, levels = unique(pdat$Species), ordered = T)
# names(pdat)
# [1] "Species"          "Sex_sys"          "Avg_stig_sp"      "Sd_stig_sp"       "n_stig"           "SE_stig_sp"      
#  [7] "Avg_prod_anth_sp" "Sd_prod_anth_sp"  "n_prod_anth"      "SE_prod_anth_sp" 
```

Plots

```{r plot efficiency, fig.dim = c(8, 6)}
pdat %>% 
  ggplot(aes(x=Avg_prod_anth_sp, y=Avg_stig_sp, shape=Sex_sys, color=Sex_sys) ) + 
  geom_point(size=5, alpha=0.8) + 
  geom_errorbar(aes(ymin = pmax(0,Avg_stig_sp - Sd_stig_sp), ymax = Avg_stig_sp + Sd_stig_sp), color="black") + 
  geom_errorbarh(aes(xmin = Avg_prod_anth_sp - Sd_prod_anth_sp, xmax = Avg_prod_anth_sp + Sd_prod_anth_sp), color="black") + 
  scale_x_continuous(name = expression("Pollen production per anther"), na.value=0) + 
  scale_y_continuous(name = "Pollen capture per flower", na.value = 0) + 
  guides(shape=guide_legend(title="Sex system"), color=guide_legend(title="Sex system")) + 
  scale_shape_manual(values = c(15, 16, 17),
                     labels=c("Dioecious", "Monoecious", "Hermaphroditic")) + 
  scale_color_manual(labels=c("Dioecious", "Monoecious", "Hermaphroditic"),
                     values=c("#66C2A5","#8DA0CB","#FC8D62")) + 
  # geom_text(aes(label=Species), size=2, color="black") + 
  theme_cs(font="sans", fontsize=20)

```

Tables

```{r}
# Create table of pollination efficiency

eff_tab <- pdat %>% dplyr::group_by(Sex_sys, Species) %>% 
  dplyr::summarize(Avg_stig_sp = Avg_stig_sp,
            Sd_stig_sp = Sd_stig_sp,
            Avg_prod_anth_sp = Avg_prod_anth_sp,
            Sd_prod_anth_sp = Sd_prod_anth_sp,
            polleff = 100 * Avg_stig_sp/Avg_prod_anth_sp)
print(eff_tab, width = 90)

write.csv(eff_tab, "processed-data/efficiency-table.csv", row.names = F)


```

