---
title: "PGLS pollen size by sex system"
author: "Claire Smith"
date: "2023-06-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

PGLS analysis based on tutorial by Roger Mundry from the Chapter 6 practical material from Modern Phylogenetic Comparative Methods and Their Application in Evolutionary Biology (Springer 2014). Link: https://www.mpcm-evolution.com/practice/online-practical-material-chapter-6/chapter-6-1exercises-testing-assumptions-statistical-issues-framework-phylogenetic-generalized-least-squares

I'll run two analyses: 

B. Does pollen size vary in wind-pollinated species according to their sex system? Specifically, do dioecious species make smaller pollen (to facilitate long-distance dispersal)? 

model: pollen size ~ sex system + collector

A. Is there an effect of sex system on pollen number? Do we see a size-number tradeoff across species (i.e. is there an effect of pollen size on number)?

model: pollen production/anther ~ pollen size + sex system + collector

Variables:
- pollen production/anther is average pollen produced per anther per species
- pollen size is the mean diameter of pollen grains per species
- sex system is a categorical test predictor with 3 levels (dioecious, monoecious, hermaphroditic)
- collector is a categorical control predictor with 2 levels (CS, JF), to control for any effects on pollen load depending on who collected the data

```{r initialize, message=F}
# Load packages
library(dplyr)
library(car)
library(V.PhyloMaker2)
library(ape)
library(dplyr)
library(caper)

library(lme4)
library(lmerTest)
library(ggplot2)
library(emmeans)

source("theme_cs.R")
```

### Read in and prepare data

Read in the tree data:

```{r read tree data}
#read the phylogeny into an R object
sizenum_tree <- read.tree("processed-data/sizenum_tree.nwk")
#inspect the tree
# str(sizenum_tree)
```

The tree has 29 tips and 27 internal nodes. Plot the tree:

```{r plot tree}
plot.phylo(x=sizenum_tree, cex=0.7, direction="upwards", no.margin=TRUE)
```

Read in and prepare trait data: 

```{r read trait data}
## read in data with pollen size/number 
sizenum <- read.csv("processed-data/size-prod-norep.csv", stringsAsFactors = T)
# head(sizenum)
# summary(sizenum)
# str(sizenum)

```

I want the data to be at the species level -- I'll take species-wide averages. I'll also filter the data so that I only keep species with at least 5 individuals. At the same time, I'll edit some species names to match those in the tree dataset (in later analyses mismatches come up as dropped data when I try to join tree and trait data together). I'll add a new column, "Species_", that has an underscore between genus and species, to match the species name format in the tree data. And I'll also add a "collector" column that specifies whether I (CS) or JF collected the data, in case it has an effect. 

```{r prepare trait data}
# species-wide averages
sizenum_sp <- sizenum %>% group_by(Sex_sys, Species, source) %>% 
  filter(!is.na(Avg_diam) & !is.na(Avg_pol_anth)) %>% 
  summarize(Avg_pol_anth = mean(Avg_pol_anth),
            Avg_diam = mean(Avg_diam),
            N = n()) %>%
  filter(N>=3) %>% # Keep only species with at least 5 individuals
  droplevels() %>% 
  # Update species names to match tree
  # Agropyron trachycaulum => Elymus trachycaulus
  # Elymus innovatus => Leymus innovatus
  mutate(Species = gsub("Agropyron trachycaulum", "Elymus trachycaulus", Species),
         Species = gsub("Elymus innovatus", "Leymus innovatus", Species)) %>% 
  mutate(Species = gsub("Bcarianatus", "Bromus carinatus", Species),
         Species = gsub("Astolonifera", "Agrostis stolonifera", Species)) %>% 
  # Species_ to match style of species names in phylogeny
  mutate(Species_ = gsub(" ", "_", Species)) %>% 
  # create column "Collector" describing who took the data
  mutate(collector = as.factor(case_when(source == "CS2021" ~ "CS",
                               source == "JF2001" ~ "JF",
                               source == "JF2004" ~ "JF"))) %>% 
  # Arrange species in order of sex system and alphabetically by species
  mutate(Sex_sys = as.character(Sex_sys),
         Sex_sys = factor(Sex_sys, levels=c("dioecious", "monoecious", "hermaphroditic")) ) %>% 
  arrange(Sex_sys, Species) %>% 
  as.data.frame()

# add this to species-level data 
# head(sizenum_sp)
# summary(sizenum_sp) # 27 species
# str(sizenum_sp) 
# sum(is.na(sizenum_sp)) # no missing data
# sample size per sex system: 
sizenum_sp %>% group_by(Sex_sys, Species) %>% summarize(n=n()) %>% group_by(Sex_sys) %>%  summarize(n=n())
# sample size per species:
print(sizenum_sp %>% group_by(Sex_sys, Species) %>% summarize(N=N), n=Inf)

```

The resulting dataset has 27 species.

Key variables:

* "Avg_pol_anth" - mean pollen production per anther, per species
* "Avg_diam" - the mean pollen diameter (per flower) per species in nm, averaged within individuals then across individuals to get a species-wide mean
* "collector" - a factor with 2 levels to control for who collected the data (CS or JF)
* "Sex_sys" - a factor with 3 levels specifying a species' sex system (dioecious, monoecious, or hermaphroditic). 

## Pre model-fitting checks:

#### Histograms and boxplots to check distributions of predictors and response variables

Pollen size: 

```{r pollen size pre fit dist}
# pollen size: 
ggplot(sizenum_sp) + 
  geom_histogram(aes(x=Avg_diam)) + 
  theme_cs()
ggplot(sizenum_sp) + 
  geom_histogram(aes(x=Avg_diam)) + 
  facet_grid(Sex_sys~.) + 
  theme_cs()
ggplot(sizenum_sp) + 
  geom_boxplot(aes(y=Avg_diam, x=Sex_sys)) + 
  xlab("Sex system") + 
  ylab("Pollen diameter") + 
  theme_cs()
#looks okay, not too skewed, no obvious outliers
```

Pollen production: 

```{r pollen production pre fit dist}
# pollen number: 
ggplot(sizenum_sp) + 
  geom_histogram(aes(x=Avg_pol_anth)) + 
  theme_cs()
# looks pretty right-skewed
ggplot(sizenum_sp) + 
  geom_histogram(aes(x=Avg_pol_anth)) + 
  facet_grid(Sex_sys~.) + 
  theme_cs()
ggplot(sizenum_sp) + 
  geom_boxplot(aes(y=Avg_pol_anth, x=Sex_sys)) + 
  xlab("Sex system") + 
  ylab("Pollen diameter") + 
  theme_cs()
# log pollen number: 
# min(sizenum_sp$Avg_pol_anth) # non-zero - can log-transform
sizenum_sp$Log_avg_pol_anth <- log(sizenum_sp$Avg_pol_anth)
ggplot(sizenum_sp) + 
  geom_histogram(aes(x=Log_avg_pol_anth)) + 
  theme_cs()
# looks more symmetrical!
ggplot(sizenum_sp) + 
  geom_histogram(aes(x=Log_avg_pol_anth)) + 
  facet_grid(Sex_sys~.) + 
  theme_cs()
ggplot(sizenum_sp) + 
  geom_boxplot(aes(y=Log_avg_pol_anth, x=Sex_sys)) + 
  xlab("Sex system") + 
  ylab("Log pollen production") + 
  theme_cs()
```

It's a little had to tell because there are only 2 species in the dioecious group, but it looks like overall pollen production tends to have a right-skewed distribution, so I will log transform it to make the distribution more symmetric. 

Sex system and collector: 

```{r sex system pre fit dist}
#inspect the frequency distributions of the factors
table(sizenum_sp$collector) #CS:1, JF:26 #only 1 CS! Will not consider collector for the analyses then
table(sizenum_sp$Sex_sys)
     # dioecious     monoecious hermaphroditic 
     #         2              8             17 
```

#### Multicollinearity 

Are there collinear predictors? Test for multicollinearity using function vif from package "car" (Fox & Weisberg 2011).

```{r test for multicollinearity}
# are size and sex system collinear? 
lm.res3 <- lm(Log_avg_pol_anth ~ Avg_diam + Sex_sys, data=sizenum_sp)
vif(mod=lm.res3) # use vif() from "car"
# no! (not over 10)
```


## Fitting models

#### Combine data

To do the phylogenetic least squares analysis (PGLS), the function pgls() needs the data and the tree to be combined into a single object, do this with the function comparative.data() from "caper":

```{r combine trait and tree data}
test.data=comparative.data(phy=sizenum_tree, data=sizenum_sp, names.col=Species_, vcv=TRUE, na.omit=FALSE, warn.dropped=TRUE)
#were any species dropped?
test.data$dropped # "Setaria_viridis"
```

### B. Effect of sex system on pollen size 

Does pollen size vary in wind-pollinated species according to their sex system? Specifically, do dioecious species make smaller pollen (to facilitate long-distance dispersal) and hermaphroditic species make larger pollen? 

First, I'll run a PGLS model where I predict the optimal branch length parameter lambda using maximum-likelihood, to test if there is phylogenetic signal in the data. 

```{r ML lambda PGLS size}
modB <- pgls(Avg_diam ~ Sex_sys, dat=test.data, lambda="ML")
summary(modB) # estimated lambda is 0.793, but not significantly different from 0 or 1
anova(modB) # no sig effect of sex system
plot(pgls.profile(modB, "lambda")) # pretty flat likelihood surface, small local peak near 0, another peak at what must be 0.793
```

The ML lambda was estimated to be 0.793, but wasn't significantly different from 0 or 1. Because of this, and because the dataset is fairly small (<30 spp), I'll run two analyses: a PGLS model with lambda=1 and an OLS model (lambda=0). I'll also run an additional mixed effects model with species as a random effect, using individual-level data rather than the species-level data used for the PGLS analysis. 

#### Pollen size PGLS model lambda=1

Fit the PGLS model with lambda=1. 

```{r PGLS size}
modB_PGLS <- pgls(Avg_diam ~ Sex_sys, dat=test.data, lambda=1)
summary(modB_PGLS)
anova(modB_PGLS) # no sig effect of sex system
```

Using a PGLS model with lambda=1, there is no significant effect of sex system on pollen size. 

Check post-fit assumptions:

```{r PGLS size post fit assumptions}
# test PGLS assumptions post fit
#inspect the phylogenetic residuals for normality:
hist(modB_PGLS$phyres) #looks pretty normal
qqnorm(modB_PGLS$phyres) 
qqline(modB_PGLS$phyres) #looks okay...
#inspect phylogenetic residuals for homogeneity:
plot(x=fitted(modB_PGLS), y=modB_PGLS$phyres, pch=19)
# only 2 dioecious species - might give them undue influence and not represent their sex system well. Also much more variable in hermaphroditic category. 
```

Residuals look relatively normal but not homoskedastic among sex system categories. This is probably mostly due to the fact that there are only 2 dioecious species, so they are necessarily less variable than the other sex systems. There only being 2 dioecious species may give these points undue influence and may mean this isn't a great representation of that sex system, which could reduce the validity of these models. 

#### Pollen size OLS model (lambda=0). 

Fit ordinary least squares model size ~ sex system. 

```{r OLS size}
modB_OLS <- lm(Avg_diam ~ Sex_sys, dat=sizenum_sp)
summary(modB_OLS)
anova(modB_OLS) # sig effect of sex system
```

In the OLS model, I found a significant effect of sex system on pollen size.

Check assumptions of the OLS model. 

```{r PGLS size post fit assumptions}
#normality of residuals
hist(modB_OLS$residuals) #looks narrow but ok
qqnorm(modB_OLS$residuals) #oh my!
qqline(modB_OLS$residuals)
#homogeneity of residuals
plot(x=fitted(modB_OLS), y=modB_OLS$residuals, pch=19) #potentially an issue here with so few dioecious species. The two dioecious species have much lower fitted values, appear to be less variable (but hard to say because there's only 2!)
```

The residuals in the qqplot diverge from the expected values a lot -- an OLS model like this might not be appropriate...

I'll also do a linear mixed effects model on pollen size by sex system with species as random effect. I'll use individual-level data instead of the species-level data necessary for the PGLS. 

```{r size mixed mod}
# Prepare individual-level data
sizenum_filt <- sizenum %>% 
  # Check all rows have both size and number measures
  filter(!is.na(Avg_diam) & !is.na(Avg_pol_anth)) %>%  
  # Make new column, collector
  mutate(collector = ifelse(source == "CS2021", "CS", "JF")) %>% 
  # Arrange species in order of sex system and alphabetically by species
  mutate(Sex_sys = as.character(Sex_sys),
         Sex_sys = factor(Sex_sys, levels=c("dioecious", "monoecious", "hermaphroditic")) ) %>% 
  arrange(desc(Sex_sys), desc(Species)) %>% 
  mutate(Species =  factor(Species, levels = unique(Species), ordered = T)) 

# sample size per species
print(sizenum_filt %>% group_by(Sex_sys, Species) %>% summarize(n=n()), n=Inf) 
# keep species with at least 10 individuals per species
spp_over10 <- sizenum_filt %>% group_by(Sex_sys, Species) %>% summarize(n=n()) %>% 
  filter(n>=10)
keep_spp <- spp_over10$Species

sizenum_filt %>% group_by(Sex_sys, Species) %>% summarize(n=n()) %>% 
  filter(Species %in% keep_spp) %>% 
  group_by(Sex_sys) %>%  summarize(n=n())

sizenum_filt2 <- sizenum_filt %>% 
  filter(Species %in% keep_spp)

size_mixmod <- lmer(Avg_diam ~ Sex_sys + (1|Species), data = sizenum_filt)
summary(size_mixmod)
anova(size_mixmod)

size_mixmod2 <- lmer(Avg_diam ~ Sex_sys + (1|Species), data = sizenum_filt2)
summary(size_mixmod2)
anova(size_mixmod2)
```

From the mixed effects model, sex system has a significant effect on pollen size. I'll use emmeans to run a Tukey's HSD to determine which groups are significantly different from each other and the direction of the difference(s). 

```{r size mixed mod emmeans}
emmeans(size_mixmod , ~Sex_sys)
emmeans(size_mixmod , list(pairwise ~ Sex_sys), adjust = "tukey")
# plot(emmeans(size_mixmod , ~Sex_sys))
em_size <- emmeans(size_mixmod , ~Sex_sys) %>%  as.data.frame()
# Plot estimated marginal means: 
ggplot(data=em_size, aes(x=Sex_sys, y=emmean)) + 
  geom_point(size=3) + 
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.1) +
  # Add letters corresponding to Tukey's HSD
  geom_text(aes(x=1, y=25), label="A") + 
  geom_text(aes(x=2, y=27.5), label="AB") + 
  geom_text(aes(x=3, y=31), label="B") + 
  scale_y_continuous(name = expression("Pollen diameter ("*mu*"m)")) + 
  scale_x_discrete(name = "Sex system", labels = c("Dioecious", "Monoecious", "Hermaphroditic")) + 
  theme_cs()
```

In the mixed effects model, sex system does have a significant effect on size. A Tukey's HSD test shows that dieocious speces have significantly smaller pollen than hermaphroditic species but not monoecious species, and monoecious species don't have significantly smaller pollen than hermaphroditic species. 

#### A. Is there an effect of sex system on pollen number? Do we see a size-number tradeoff across species (i.e. is there an effect of pollen size on number)?

model: pollen production/anther ~ pollen size + sex system + collector

```{r PGLS model C prod by size (tradeoff) and sex system}
# fit the PGLS model:
fullA <- pgls(Log_avg_pol_anth ~ Avg_diam + Sex_sys, dat=test.data, lambda="ML")
summary(fullA)
# ML lambda estimate is 0, not significantly different from 0 or 1. 
anova(fullA) # sig effect of diam and sex sys
plot(pgls.profile(fullA, "lambda")) # strange likelihood surface, looks like local peaks around 0 and 1...
# Since the ML lambda wasn't significantly different from 0 or 1 and the dataset is fairly small (<30 spp), I'll run
# two analyses: an OLS model (lambda=0) and a PGLS model with lambda=1
# OLS: 
modA_OLS <- lm(Log_avg_pol_anth ~ Avg_diam + Sex_sys + collector, dat=sizenum_sp)
summary(modA_OLS)
anova(modA_OLS) # sig effect of sex system and diameter

# PGLS with lambda=1:
modA_PGLS <- pgls(Log_avg_pol_anth ~ Avg_diam + Sex_sys + collector, dat=test.data, lambda=1)
summary(modA_PGLS)
anova(modA_PGLS) # no sig effect of sex system or diameter

### Check assumptions: 
## OLS
#normality of residuals
hist(modA_OLS$residuals) #ok
qqnorm(modA_OLS$residuals)
qqline(modA_OLS$residuals) #looks okay
#homogeneity of residuals
plot(x=fitted(modA_OLS), y=modA_OLS$residuals, pch=19) 

##PGLS
#inspect the phylogenetic residuals for normality:
hist(modA_PGLS$phyres) #looks pretty normal
qqnorm(modA_PGLS$phyres) 
qqline(modA_PGLS$phyres) #looks okay...
#inspect phylogenetic residuals for homogeneity:
plot(x=fitted(modA_PGLS), y=modA_PGLS$phyres, pch=19)
# only 2 dieocious species - might give them undue influence and not represent their sex system well. Also much 
# more variable in hermaphroditic category. 

## Mixed effects mod
# Since lambda not significantly different from 0 or 1, I'll also do an lmer OLS model with species as random effect.
# I'll use the data that's not grouped by species - sizenum
# head(sizenum)
# summary(sizenum)
# names(sizenum)
#  [1] "source"          "Sex_sys"         "Species"         "Site"            "Ind"             "Label"          
#  [7] "Avg_diam"        "Sd_diam"         "N_diam"          "Avg_area"        "Avg_pol_anth"    "Sd_pol_anth"    
# [13] "Anth_per_flw"    "count_method"    "Anth_per_sample" "Pol_flw"         "Sd_pol_flw"
```

```{r}
sizenum_filt <- sizenum %>% 
  # Check all have both size and number measures
  filter(!is.na(Avg_diam) & !is.na(Avg_pol_anth)) %>%  
  filter(Species %in% sizenum_sp$Species) %>% 
  # Make new column, collector
  mutate(collector = ifelse(source == "CS2021", "CS", "JF")) %>% 
  # Create Log pol anth column
  mutate(Log_avg_pol_anth = log(Avg_pol_anth))

prodsize_mixmod <- lmer(Log_avg_pol_anth ~ Avg_diam + Sex_sys + (1|Species), data = sizenum_filt)
summary(prodsize_mixmod)
anova(prodsize_mixmod) # sig effects of diam and sex system
vif(prodsize_mixmod)

emmeans(prodsize_mixmod , ~Sex_sys)
emmeans(prodsize_mixmod , list(pairwise ~ Sex_sys), adjust = "tukey")
# plot(emmeans(prodsize_mixmod , ~Sex_sys))
em_prodsize <- emmeans(prodsize_mixmod , ~Sex_sys) %>%  as.data.frame()
# Plot estimated marginal means for pollen production: 
ggplot(data=em_prodsize, aes(x=Sex_sys, y=emmean)) + 
  geom_point(size=3) + 
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.1) +
  # Add letters corresponding to Tukey's HSD
  geom_text(aes(x=1, y=10.1), label="AB") + 
  geom_text(aes(x=2, y=9.6), label="A") + 
  geom_text(aes(x=3, y=8.1), label="B") + 
  scale_y_continuous(name="Pollen production per anther",
                     breaks = log(c(250, 500,1000, 2500, 5000, 10000, 20000)),
                     labels = c(250, 500,1000, 2500, 5000, 10000, 20000),
                     limits = c(log(250), NA)) +
  scale_x_discrete(name = "Sex system", labels = c("Dioecious", "Monoecious", "Hermaphroditic")) + 
  theme_cs()
```

Hermaphroditic species produce significantly more pollen than monoecious species, when size and sex system are
accounted for. Pollen production does not differ significantly between monoecious and dioecious species or between dioecious and hermaphroditic species. 

Plot pollen number by size to visualize tradeoff. 
```{r}
emmip(prodsize_mixmod ,  ~ Avg_diam, adjust = "tukey")
emms <- emmip(prodsize_mixmod, ~ Avg_diam, cov.reduce=range,
              type="response", plotit=F)
em_prodsize <- emmeans(prodsize_mixmod , ~Sex_sys) %>%  as.data.frame()
# Plot estimated marginal means for pollen production: 
ggplot(data=em_prodsize, aes(x=Sex_sys, y=emmean)) + 
  geom_point(size=3) + 
  geom_errorbar(aes(ymin=lower.CL, ymax=upper.CL), width=0.1) +
  # Add letters corresponding to Tukey's HSD
  geom_text(aes(x=1, y=10.1), label="AB") + 
  geom_text(aes(x=2, y=9.6), label="A") + 
  geom_text(aes(x=3, y=8.1), label="B") + 
  scale_y_continuous(name="Pollen production per anther",
                     breaks = log(c(250, 500,1000, 2500, 5000, 10000, 20000)),
                     labels = c(250, 500,1000, 2500, 5000, 10000, 20000),
                     limits = c(log(250), NA)) +
  scale_x_discrete(name = "Sex system", labels = c("Dioecious", "Monoecious", "Hermaphroditic")) + 
  theme_cs()
```


