---
title: "Stigma pollen by sex sys PGLS"
author: "Claire Smith"
date: "2023-06-27"
output: github_document
editor_options: 
  chunk_output_type: console
---

Analysing the relationship between average stigmatic pollen loads and sex system across 22 wind-pollinated flowering plant species. 

```{r initialize, message=F}
# Load packages
library(tidyverse)
library(car)
library(V.PhyloMaker2)
library(ape)
library(caper)
library(knitr)
library(lme4)
library(emmeans)

source("theme_cs.R")
```

#### Question

Is there an effect of sex system on pollen load, across species? 

stigmatic pollen load ~ sex system + collector

- stigmatic pollen load is the response variable (continuous numeric), the mean pollen load per stigma per species 
- sex system is a categorical test predictor with 3 levels (dioecious, monoecious, hermaphroditic)
- collector is a categorical control predictor with 2 levels (CS, JF), to control for any effects on pollen load depending on who collected the data

```{r prepare tree data}
#read the phylogeny into an R object
stig_tree <- read.tree("processed-data/capture_tree.nwk")
#inspect the tree
str(stig_tree)
#sn.tree is a tree with 22 tips and 20 internal nodes
#plot the tree:
plot.phylo(x=stig_tree, cex=0.7, direction="upwards", no.margin=TRUE)
```

```{r prepare trait data}
## read in data with stigmatic pollen capture and potential covariates
stig <- read.csv("processed-data/stig-no-rep-spp.csv", stringsAsFactors = T) 
# head(stig)
# summary(stig)

# I want the data to be at the species level -- I'll take species-wide averages
stig_sp <- stig %>% group_by(Sex_sys, Species, source) %>% 
  filter(!is.na(Flw_pollen)) %>% 
  summarize(mean_poll = mean(Flw_pollen))
# head(stig_sp)
# summary(stig_sp)
# str(stig_sp)
# View(stig_sp)

# Add some helpful columns
stig_sp <- stig_sp %>% 
  # Correct the spelling of "Schizachne purpurascens"
  mutate(Species = gsub("Schizacne purpurascens", "Schizachne purpurascens", Species)) %>% 
  # Update "Elymus innovatus" to "Leymus innovatus"
  mutate(Species = gsub("Elymus innovatus", "Leymus innovatus", Species)) %>% 
  # create column "Species_" so that species names in this dataset match tip labels in the tree
  mutate(Species_ = gsub(" ", "_", Species)) %>% 
  # create column "Collector" describing who took the data
  mutate(collector = as.factor(case_when(source == "CS2021" ~ "CS",
                               source == "JF2001" ~ "JF",
                               source == "JF2004" ~ "JF"))) %>% 
  as.data.frame()

# Inspect
# str(stig_sp) # 23 species 
# "mean_poll" - mean stigmatic pollen load per species
# "collector" - a factor with 2 levels to control for who collected the data (CS or JF)
# "Sex_sys" - a factor with 3 levels specifying a species' sex system (dioecious, monoecious, or hermaphroditic). 
# check whether there are any missing values:
# sum(is.na(stig_sp)) # no missing values
```

Pre model-fitting checks:

```{r inspect distributions, include=F}
#frequency distributions of categorical predictors
table(stig_sp$collector) #CS is represented by 10 species, JF is represented by 13
table(stig_sp$Sex_sys) #there are only 2 dioecious species, but 12 hermaphroditic species and 9 monoecious species. This might be an issue - the dioecious species may be overly influential and also not a great representation of this sex system. 

#inspect the distribution of the response:
hist(stig_sp$mean_poll) #this is very skewed!
# I'll check whether this looks more symmetrical after a log-transformation (to avoid influential cases)
#first I'll check if this contains any 0s...
min(stig_sp$mean_poll) #it's 2.26... which is >0 so a simple log-transform is safe
hist(log(stig_sp$mean_poll)) #looks better (not heavily skewed, no outliers)
stig_sp$log.mean_poll <- log(stig_sp$mean_poll) #add log-transformed response to dataset
```


```{r checking for collinearity, include=F}
#run a linear model wwith all predictors
lm.res <- lm(log.mean_poll ~ Sex_sys + collector, data=stig_sp)
vif(mod=lm.res) # variance inflation factors from package "car"
#generalized inflation factors are pretty small (under 10) -- probably no big collinearity issues
```

Model fitting: 

Join trait and tree data: 

```{r fit ML PGLS model}
#the function pgls() needs the data and the tree to be combined into a single object, do this with the function comparative.data() from "caper"
# note!!!: trait data CANNOT be a tibble, make sure to convert it to a dataframe or you will get an error!
test.data=comparative.data(phy=stig_tree, data=stig_sp, names.col=Species_, vcv=TRUE, na.omit=FALSE, warn.dropped=TRUE)
# Check for mismatched names:
# stig_tree$tip.label[!(stig_tree$tip.label %in% stig_sp$Species_)]
# -> fixed spelling of Schizachne purpurascens, updated Elymus innovatus to Leymus innovatus

#were any species dropped?
# test.data$dropped # no!
```

Fit PGLS model 

```{r}
#now fit the PGLS model with maximum likelihood estimated lambda
# full <- pgls(log.mean_poll ~ Sex_sys + collector, dat=test.data, lambda="ML")
# #the argument lambda="ML" means that lambda, the scaling parameter of the variance-covariance matrix is estimated (using maximum likelihood)
# summary(full)
# anova(full)

#drop collector since it's not significant
full2 <- pgls(log.mean_poll ~ Sex_sys, dat=test.data, lambda="ML")
summary(full2)
# anova(full2)
kable(anova(full2))

# PGLS branch length output from summary()
# Branch length transformations:
# kappa  [Fix]  : 1.000
# lambda [ ML]  : 0.738
#    lower bound : 0.000, p = 0.50889
#    upper bound : 1.000, p = 0.0046432
#    95.0% CI   : (NA, 0.977)
# delta  [Fix]  : 1.000
# From the above - the branch scaling parameter lambda isn't significantly different from 0 but it is significantly different from 1. Lambda=0 is equivalent to an OLS model/a "star" phylogeny where no species are more related to each other than others. Lambda=1 represents a Brownian model of evolution. In terms of this model, might mean that PGLS model is no more appropriate than an OLS model because the differences in average pollen load between species aren't particularly affected by where they sit on a phylogeny. 
# I will take a look at lambda's likelihood profile: 
plot(pgls.profile(full2, "lambda"))
```

Lambda is significantly different from 1 but not 0 - I'll run two models, one with the ML lambda (lambda=0.738) and another with lambda=0, equivalent to an OLS model. 

Fitting OLS model and PGLS model with lambda=0.738:

```{r fit final models}
#OLS model 
# OLS_stig <- lm(log.mean_poll ~ Sex_sys + collector, dat=stig_sp)
# summary(OLS_stig)
# anova(OLS_stig) # again, I'll remove collector since it's not significant and I didn't expect it to have an impact
OLS_stig <- lm(log.mean_poll ~ Sex_sys, dat=stig_sp)
summary(OLS_stig)
anova(OLS_stig)

#PGLS model with lambda=1
PGLS_stig <- pgls(log.mean_poll ~ Sex_sys, dat=test.data, lambda="ML")
summary(PGLS_stig)
anova(PGLS_stig)
```

When using an OLS model, sex system has a significant effect on log mean pollen load (F20,2 = 9.22, p=0.0015). However, using the PGLS model with lambda=0.738 this effect is absent (F19,2=0.80, p=0.4636). 

Post model-fitting checks:

```{r check residuals, include=F}
## OLS
#normality of residuals
hist(OLS_stig$residuals) #looks ok
qqnorm(OLS_stig$residuals) 
qqline(OLS_stig$residuals) #looks okay
#homogeneity of residuals
plot(x=fitted(OLS_stig), y=OLS_stig$residuals, pch=19) #potentially an issue here with so few dioecious species. The two dioecious species have much lower fitted values, appear to be less variable (but hard to say because there's only 2!)

#PGLS
#normality of phylogenetic residuals:
hist(PGLS_stig$phyres) #not bad
qqnorm(PGLS_stig$phyres) 
qqline(PGLS_stig$phyres) #look ok
#inspect phylogenetic residuals for homogeneity:
plot(x=fitted(PGLS_stig), y=PGLS_stig$phyres, pch=19)
#again -- two dioecious species have much lower fitted values, appear to be less variable (but hard to say because there's only 2!)
```

Since lambda was not significantly different from 0, I'll also do a linear mixed effects model on pollen receipt by sex system with species as random effect.

```{r stig mixed mod}
## Mixed effects mod
# I'll use the data that's not grouped by species - sizenum
stig_filt <- stig %>% 
  filter(!is.na(Flw_pollen)) %>% 
  # Make log mean poll column %>% 
  mutate(Log_avg_poll = log(Flw_pollen + 1)) %>% 
  # Make new column, collector
  mutate(collector = ifelse(source == "CS2021", "CS", "JF")) %>% 
  # Arrange species in order of sex system and alphabetically by species
  mutate(Sex_sys = as.character(Sex_sys),
         Sex_sys = factor(Sex_sys, levels=c("dioecious", "monoecious", "hermaphroditic")) ) %>% 
  arrange(desc(Sex_sys), desc(Species)) %>% 
  mutate(Species =  factor(Species, levels = unique(Species), ordered = T)) 

stig_mixmod <- lmer(Log_avg_poll ~ Sex_sys + (1|Species/Plant), data = stig_filt)
summary(stig_mixmod)
# anova(stig_mixmod)
kable(anova(stig_mixmod))
```

From the mixed effects model, sex system has a significant effect on stigmatic pollen load. I'll use emmeans to run a Tukey's HSD to determine which groups are significantly different from each other and the direction of the difference(s). 

```{r stig mixed mod emmeans}
emmeans(stig_mixmod , ~Sex_sys)
emmeans(stig_mixmod , list(pairwise ~ Sex_sys), adjust = "tukey")
# plot(emmeans(size_mixmod , ~Sex_sys))
em_stig <- emmeans(stig_mixmod , ~Sex_sys) %>%  as.data.frame()
```

Plot estimated marginal means for stigma load by sex system 

```{r plot estimated marginal means stig load, dpi=200, fig.dim=c(12,12)}
# Plot estimated marginal means: 
ggplot(data=em_stig, aes(x=Sex_sys, y=emmean)) + 
  geom_point(size=5) + 
  geom_errorbar(aes(ymin=pmax(0,asymp.LCL), ymax=asymp.UCL), width=0.1) +
  # Add letters corresponding to Tukey's HSD
  geom_text(aes(x=1, y=log(10)), label="A", size=6) + 
  geom_text(aes(x=2, y=log(13)), label="A", size=6) + 
  geom_text(aes(x=3, y=log(42)), label="B", size=6) + 
  scale_y_continuous(name="Stigmatic pollen load",
                     breaks = log(c(0,1,2,5,10,25,50,100,200,500,1000, 2500)+1),
                     labels = c(0,1,2,5,10,25,50,100,200,500,1000, 2500),
                     limits=c(0,NA)) +
  scale_x_discrete(name = "Sex system", labels = c("Dioecious", "Monoecious", "Hermaphroditic")) + 
  theme_cs(fontsize = 20)
```

In the mixed effects model, sex system does have a significant effect on stigmatic pollen loads. A Tukey's HSD test shows that hermaphroditic species capture significantly more pollen than dioecious or monoecious species, but pollen capture is not significantly different between dioecious and monoecious species. 

Plot estimated marginal means for pollen size per sex system, with points for means: 

```{r plot emms size sex sys with points, dpi=200, fig.dim=c(5,5)}
stig_ind <- stig_filt %>% group_by(Species, Sex_sys, Date, Site, Plant) %>% 
  summarize(mean_poll=mean(Flw_pollen, na.rm=T),
            log.mean_poll=log(mean_poll+1))

# Plot estimated marginal means: 
ggplot(data=em_stig, aes(x=Sex_sys, y=emmean)) + 
  geom_point(size=3) + 
  geom_errorbar(aes(ymin=pmax(0,asymp.LCL), ymax=asymp.UCL), width=0.1) +
    # Add letters corresponding to Tukey's HSD
  geom_text(aes(x=1, y=log(13)), label="A", size=6) + 
  geom_text(aes(x=2, y=log(17)), label="A", size=6) + 
  geom_text(aes(x=3, y=log(53)), label="B", size=6) + 

  # Raw data (individual means) with stat dot histograms
  # ggdist::stat_dots(data=sizenum, aes(x=Sex_sys, y=Avg_diam),
  #   side = "left",
  #   dotsize = .8,
  #   justification = 1.2,
  #   binwidth = .3) +
  # Raw data (individual means) with jittered points
  gghalves::geom_half_point(data=stig_ind, aes(x=Sex_sys, y=log.mean_poll),
    ## draw jitter on the left
    side = "l",
    ## control range of jitter
    range_scale = .4,
    ## add some transparency
    alpha = .3,
    size=1) +
#   
#   # Species means
  gghalves::geom_half_point(data=stig_sp, aes(x=Sex_sys, y=log.mean_poll),
    ## draw jitter on the right
    side = "r",
    ## control range of jitter
    transformation = position_identity(),
    # transformation = position_jitter(w = 0, h = 0),  # we do NOT want vertical jitter!
    ## add some transparency
    alpha = .7,
    size=5,
    shape=95) +
  scale_y_continuous(name="Stigmatic pollen load",
                     breaks = log(c(0,1,2,5,10,25,50,100,200,500,1000, 2500)+1),
                     labels = c(0,1,2,5,10,25,50,100,200,500,1000, 2500),
                     limits=c(0,NA)) +
  scale_x_discrete(name = "Sex system", labels = c("Dioecious", "Monoecious", "Hermaphroditic")) + 
  theme_cs()
```
